\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-03 syntax.tex
Small fixes throughout metatheory section

We formalize our language as a deep embedding in a type theory with QIIT types. Every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms. (Otherwise how can they be encoded in Agda?) There are four standard judgement forms in MLTT,
well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. These four judgements (QIIT types) are
mutually recursively defined together. Our presentation
follows the style of~\citet{kaposi2019gluing}, but for readability we
omit the universe levels. We still working on predicative MLTT.
%\YZ{What's the distinction of judgments vs. sequents?}.\EDJreply{I am not sure about the correct terminology here. Here I want to say "even underivable judgement has to be well-typed". That's the main reason I spell out the fake Agda code there. I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}\EDJreply{I rephrase that so there won't be confusion for the future readers. Please check}
Latex-wise, we spell out these \textit{well-formedness rules} of type
judgements and their Agda counterparts to make things more accessible. Of
course, a well-formed judgement is not necessarily derivable, just as
a well-formed QIIT type is not necessarily inhabited.

Due to our intrinsic setting, \textbf{we can and will omit a lot of
presumptions of the typing rules}. This can work because we are using
QIIT and everything is intrinsically well-formed. For example, we don't
spell out that the rule \ruleref{Type Universe} requires a well-formed
context, even though it does implicitly, \textbf{because it is not
possible to have ill-formed contexts in our intrinsic setting}. This is
a bit unusual for conventional extrinsic setting where they will put
well-formedness of context everywhere in a lot of term judgements. We
consider this as one advantage of using intrinsic setting---this
well-formed context is usually attached to some other presumptions thus
inferrable and can be considered as ``boilerplate code''. Free of
such boilerplate code, our intrinsic setting provides a more concise
representation.

@@

Now we highlight some other special features that differ from a more
conventional setting: (1) This modern formulation of type theory does not
have an operational semantics but only equality.\footnote{Later we will prove
canonicity which can imply a certain level of computational ability of
our theory}
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
(2) Instead of using meta-level substitution, we use explicit
substitution and de Bruijn indices in our intrinsic formulation. This is
also called substitution calculus in the literature and is favoured due
to its closedness categorical semantic.
(3) We are dealing with quotient data, so we are actually manipulating
equivalence classes of terms.

We quickly review the substitution calculus.
Substitution is very (if not the most) important because function is
first-class citizens, and can be applied (almost) everywhere and can return
(almost) everything, thus we require the resulting substitution to
commute with most (if not all) type/term constructs.

A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
understood in the way that $\gamma$ will output a \textbf{series} of
(dependently-typed related) terms of types $\Delta$, parametrized by
$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
Thus (1) \ruleref{Sub Extend} looks like concatenating one
series of terms with a new term (the weird looking type of
$\goodTerm{\Gamma}{t}{A[\sigma]}$ is because we need $A$ to be typed in
the context $\Delta$ so that we can use \ruleref{Context Extension} to
create new context); (2) we have \ruleref{Proj Subst-1} to extract a
subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
substitution really looks like a (deeply-nested) pair.

@@

$$
\quad 
\Rule[name=Func/DPair Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}
\quad \goodType{\Delta}{A}{} 
\quad \goodType{\Delta, A}{B}{}
}
{
  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
  \quad 
  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
}
$$
%\YZ{What is the requirement on Delta?}
%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context. This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well). For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing. What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda. For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer fake agda formulation) }\YZreply{I was thinking about 'Delta |- A'. I suppose it is also implicitly enforced via intrinsically typed syntax. Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right! I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU[\gamma] \equiv \cU }{j + 1} \quad
  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$
\judgebox{ \goodTerm{\Gamma}{t}{T} }
$$
\Rule[]
{\goodType{\Gamma}{T}{j}}
{\goodTerm{\Gamma}{"c" \ T}{\cU}
}\quad
\Rule[]
{\goodTerm{\Gamma}{T}{\cU}}
{\goodType{\Gamma}{El \ T}{j}}
\quad
\Rule
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad 
\Rule[name=Term Subst]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
$$
$$
\Rule[]
{\goodTerm{\Gamma, A}{t}{B}}
{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{u}{A} 
\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Sigma A B}}
{\goodTerm{\Gamma}{"pjl" \ t}{A}
\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
}
$$
$$
\Rule
{}
{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
}
$$

@@

We have shown some exemplar rules for all four standard types of
judgments. Now we focus on the newly introduced facility.

We have two kinds of signatures, one for inductive types, the other for linkages.

@@

Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A
\vdash B(x)$ decides a W-type. To simulate extensible inductive types, we
have to equip our inductive types with multiple constructors, and for
simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ :
(x : Aᵢ) → (Bᵢ x → W) → W}.
Thus, we use a list of pairs ($A_i, B_i$) as the signature of one
inductive type, and given an inductive type signature, we use
\ruleref{WSig-Proj} to extract the corresponding $A_i, B_i$.

Of course we need to put an eye on the substitution law for each piece
of the syntax. Here $\bW \sigma$ is a large singleton
type~\cite{stone2000}, and it only has one ``element'', the inductive
type itself, inside. We formulate in this cumbersome way because we want
to have inductive type to be a field element and exposing the
constructors at the type level for later \textit{pattern-matching
exhaustiveness checking} (i.e. if an inductive type as field element has type
$\cU$, then no information about the inductive type is exposed).

@@

$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
for linkages in the context $\Gamma$ with $n$ fields. We omit the
naming/label information for fields in the meta\-theory. The signature (for the linkage) and the linkage rules are quite similar to
the signature for inductive types.  They are both constructed from the
empty stuff, and respect substitution lemmas. We have
\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component
due to harmony~\cite{pfenning2009lecture}, so we have local soundness
and local completeness ($\beta,\eta$ rules). We omit them and the
substitution laws for either field addition and projection. For
$\mu^+,\nu^+$ rules, we will omit the sealing $\{f\}$ if possible.\footnote{This
omission can be supported by Agda using Implicit Arguments}

A side note: before we dive in more details of our metatheory for linkage (especially the meaning of sealing and $\cC$), we have to mention that, the challenge of incorporating family polymorphism with inductive type mentioned in \ref{chg:extensible-inductive-type} mutates in metatheory because of de Bruijn indices. (Using surface synax), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;..\}$ with a bunch of $b_i$ \textit{using} $a$. The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $a : \bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that and thus we hope $b_i$ constructed in the context of $a : \cU$ (and thus later this $a$ can be compatible with inductive type of any signature). Our metatheory formulation of linkage is mainly trying to solve this tension.

We use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).
$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, according to \ruleref{Compile}, is a deeply nested dependent pair. Then $\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.

The reason we need this additional type construct of $\cC\sigma$ here is
to support \textit{code reuse}, as the rule of linkages hinders its
ability to do abstraction. Concretely speaking, (using surface syntax)
we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to
T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one,
consider the case $T$ is bottom type and $\sigma$ is empty record and
thus $a$ is initially bottom type. Now the direct projection using
\ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will
lead to problem because every field in linkage has this implicit
universal quantification, and for this case we can substitute $a$ with
arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top
\to \bot}$ is a contradiction}.
However, this underivable term is important for us to abstract the
concrete inductive type so that the later field can depend on the
``abstract interface'' (i.e. something like $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}. 
\YZ{It reads as if you want to have 'h : L\{a : U, b : a -> T\}'. But it is problematic because it leads to inconsistency, as you describe in a footnote.}\EDJreply{Good point. I rephrased it so the introduction of sigma type is smoother. Please check.}
\\ The good news, we don't have to stick with $\cL(\sigma)$ to express the abstract interface. We can also use sigma type, and actually 
we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
}{h}{(\Sigma a : \cU, a \to T)}$ because of the rules for sigma type.\footnote{The intro rule for sigma type decides that there is no such implicit universal quantification happening.}
\YZ{This sentence still speaks of overriding. Explain in terms of projection instead?}\EDJreply{Fixed. Please check.}
Thus to allow abstraction happens, we need $\cC\sigma$ to compile a
linkage to module as a part of the context and we also need
\ruleref{Lnkg Compile} to specify the compilation from linkage to
module.

This is also reflected in the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}. Thus we need $\cC\sigma$ in our context.
%\YZ{Why is Csigma needed? How is it different from Lsigma?}\EDJreply{It corresponds to the compilation of the family ``type'', which is part of the implementation detail. Basically we know every field of a family is dependent on the former fields. The implementation detail here is that, every field of a family is dependent on the compiled module of the former family (and thus includes all the former fields)}\YZreply{I don't think you should explain the metatheory from the angle of how things are implemented in your plugin. Rather, when describing your Coq plugin, you can say it matches the metatheory.
%PS: you use two terminologies: 'compile' and 'seal', but I guess they mean very similar things. So avoid saying 'compile' here.
%PPS: Sealing means hiding, so what is being hided? Why do you need to hide it? And how does the 'Compile' and 'Seal' rules perform hiding?}\EDJreply{  Compile is very different from seal.  Compile is making a linkage into a module, depriving the ability to override---if we understand a linkage as a module with a lot of universal quantifier inside. Compile cannot hide things, but only translate a linkage into a module. Seal is really something that does hiding---because existential type can hide stuff. Seal is an term between existential types. \\
%The ultimate reason we have this two notion is that, the introduction rule for linkage fails to provide abstraction mechanism so we have to translate them into module. \\
%Generally speaking Compile : Linkage -> Module, Seal : Module -> Module, we need module here because linkage cannot do abstraction of inductive type. A concrete detail is at the above. \\ 
%I now describe the motivation of compile in the earlier paragraph, right before the "reflecting" section. Note that, this compile is inspired by plugin implementation. This is the reason I say the metatheory is really close and inspired by the implementation. Similarly as I say ``there is no extensible inductive type but only overriding of new inductive type'' because in metatheory it is using overriding to ``extend'' (same as plugin implementation), seal means the same thing, and ``compile'' in metatheory happens in plugin implementation
%}

@@

For notation clarity, we will sometimes use $"sf'"$ to emphasize
that we consider that seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]} $ and "sf" notation (used in \ruleref{Compile}) is defined as $\goodSub{\Gamma, \cC \sigma}{sf~s := (\pi_1, sf'~s)}{\Gamma, A}$.\YZ{What is sf and sf'? sf first appears in Compile rules, so explain sf there.}\EDJreply{Fixed. They both are explained at here now}

Sealing judgement is still about the important term $h$ we mentioned
earlier. Thanks to $\cC$, a field $t_{new}$ that depends on $a$ is now
in the context containing $\cC\{a : \bW \sigma; ..\}$.
But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is
extended, we would like $t_{new}$ to be in the context containing
$\cC\{a : \cU; ..\}$ instead. The sealing is responsible for this
because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma;
..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW
\sigma,..)}{s}{(\Sigma a : \cU,..)}$. Then we can construct
$t_{new}$ in the context containing $(\Sigma a : \cU,..)$, ready for
inheritance.
Apparently $t_{new}$ in this case cannot be doing \textit{exhaustiveness
checking}, but it can freely use any \textit{exhaustiveness-checked}
recursor.
%\YZ{It occurs to me that you use the word "sealing" to mean
%different (though tangentially related) things. Previously you used it
%as in "sealed families".}\EDJreply{They are actually the same thing.
%This sealing in metatheory is much stronger than that of the implemented
%plugin. Downside of the strong sealing is that if the user really use
%this version of sealing, the user will have to annotate what the sealed
%family type is. Now in the plugin, the sealed family type is almost
%fixed (directly inferred using a fixed rules). So basically this
%complete version of sealing (in metatheory) can deduce that of seal (in
%plugin implementation)}

@@

Now, together with sealing, we can read \ruleref{Sig Add} and \ruleref{Lnkg Add} out clearly: given a constructed linkage $o : \cL\sigma$ of signature $\sigma$, the next added field $t : T$ will have to be in the context of the abtraction $A$ of the compilation $\cC\sigma$ (witnessed by the seal $f$). The ``concatenation'' of the $o$ and $t$ leads to a new linkage, of which the signature is the ``concatenation'' of the $\sigma$ and $T$ (we also record the sealing $f$ for simplicity). The construction of linkage in \ruleref{Lnkg Add} requires $t$ to \textbf{only} refer to the (abstraction $A$ of) former part of the defining family $o : \cL\sigma$, this is to avoid inconsistency mentioned in \ref{chg:consistency} in metatheory.

>>>>

We formalize our language as a deep embedding in a type theory with QIIT types. Every judgement is a (dependent) QIIT \textit{type} and all judgement derivations are well-typed QIIT terms. (Otherwise how can they be encoded in Agda?) There are four standard judgement forms in MLTT,
well-typed contexts $\goodCtx{\Gamma}{l}$, well-typed types
$\goodType{\Gamma}{T}{l}$, well-typed terms $\goodTerm{\Gamma}{t}{T}$, and
well-typed substitutions $\goodSub{\Gamma}{\gamma}{\Delta}$. These four judgements (QIIT types) are
mutually recursively defined together. Our presentation
follows the style of~\citet{kaposi2019gluing} and thus a predicative MLTT, but for readability we
omit the universe levels.
%\YZ{What's the distinction of judgments vs. sequents?}.\EDJreply{I am not sure about the correct terminology here. Here I want to say "even underivable judgement has to be well-typed". That's the main reason I spell out the (pseudo) Agda code there. I actually want the reader to have intrinsic-typed syntax in mind all the time (like read conventional math but acknowledge we are manipulating Agda)}\EDJreply{I rephrase that so there won't be confusion for the future readers. Please check}
Latex-wise, we spell out these \textit{well-formedness rules} of type
judgements and their Agda counterparts to make things more accessible. Of
course, a well-formed judgement is not necessarily derivable, just as
a well-formed QIIT type is not necessarily inhabited. These well-formedness rules are direct translations of the \textit{well-typedness} of QIIT.

Due to our intrinsic setting, \textbf{we can and will omit a lot of premises of the typing rules}. This can work because we are using
QIIT and everything is intrinsically well-formed(well-typed). For example, we don't
spell out that the rule \ruleref{Type Universe} requires a well-formed
context, even though it does implicitly, \textbf{because it is not
possible to have ill-formed(ill-typed) contexts in our intrinsic setting}. This is
a bit unusual for conventional extrinsic setting where they will put
well-formedness of context everywhere in a lot of term judgements. We
consider this as one advantage of using intrinsic setting---this
``well-formed context'' condition is usually attached to some other premises thus
inferrable and can be considered as ``boilerplate code''. Free of
such boilerplate code, our intrinsic setting provides a more concise
representation.

@@

Now we highlight some other special features that differ from a more
conventional setting: (1) This modern formulation of type theory does not
have an operational semantics but only equality.\footnote{Later we will prove
canonicity which can imply a certain level of computational ability of
our theory}
%We will recover canonicity afterwards, which implies the computational ability of our theory. 
(2) Instead of using meta-level substitution, we use explicit
substitution and de Bruijn indices in our intrinsic formulation. This is
also called substitution calculus in the literature and is favoured due
to its closedness to categorical semantic.
(3) We are dealing with quotient data, so we are actually manipulating
equivalence classes of terms instead of one representative (term) from the equivalence class.

We quickly review the substitution calculus.
Substitution is very (if not the most) important because function is
first-class citizens, and can be applied (almost) everywhere and can return
(almost) everything, thus we require \textbf{the resulting substitution to
commute} with most (if not all) type/term constructs.

A term $\goodTerm{\Gamma}{t}{A}$ can be understood as one term $t$ of
type $A$, with a bunch of holes (of type $\Gamma$) inside the term $t$.
A substitution $\goodSub{\Gamma}{\gamma}{\Delta}$ can be similarly
understood in the way that $\gamma$ will output a \textbf{series} of
(dependently-typed related) terms of types $\Delta$, parametrized by
$\Gamma$ (i.e., we can consider there are holes of type $\Gamma$ in $\gamma$).
Thus (1) \ruleref{Sub Extend} looks like concatenating one
series of terms with a new term (the weird looking type of
$\goodTerm{\Gamma}{t}{A[\sigma]}$ in the premise is due to our need of type $A$ in
the context $\Delta$ so that we can use \ruleref{Context Extension} to
create new context); (2) we have \ruleref{Proj Subst-1} to extract a
subpart of a given substitution; (3) \ruleref{Proj-Ext} explains why
substitution really looks like a (deeply-nested) pair.

@@

$$
\quad 
\Rule[name=Func/DPair Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}
\quad \goodType{\Delta}{A}{} 
\quad \goodType{\Delta, A}{B}{}
}
{
  \goodType{\Gamma}{(\Pi A B)[\gamma] \equiv \Pi A[\gamma] B[\gamma^\uparrow] }{j \cup k}
  \quad 
  \goodType{\Gamma}{(\Sigma A B)[\gamma] \equiv \Sigma A[\gamma] B[\gamma^\uparrow] }{j \cup k}
}
$$
%\YZ{What is the requirement on Delta?}
%\EDJreply{ Delta here is of type Con and thus Delta here should be a well-typed Context. This is another issue I want to talk about, in Agda these kinds of stuff is inferred (so the reader can infer them as well). For example, we know delta is at the place of context, so it is a context, but we don't need to specify that it is a ``well-typed context'' because every context is well-typed in this intrinsic-typing. What's more, I think adding this extra judgement here just cause significant code bloat (like what I complained to you before I know how to use implcit variable in agda. For example, TYPE SUBST, there will be two more judgements (with no more information there) ) (Another reason I prefer (pseudo) agda formulation) }\YZreply{I was thinking about 'Delta |- A'. I suppose it is also implicitly enforced via intrinsically typed syntax. Seems to me it is worth stating, however, or otherwise the rule appears to say that Delta can be any context as long as Gamma |- gamma : Delta.}\EDJreply{You are absolutely right! I think Delta |- A, even though inferrable, mentioning it is good, also some information are B is also ok, please check, resolve if ok, thanks!}
$$
\Rule[name=Type Subst]
{\goodType{\Delta}{T}{j} 
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodType{\Gamma}{T[\gamma]}{j}}
\quad
\Rule[name=Base Type Subst]
{\goodSub{\Gamma}{\gamma}{\Delta}}
{\goodType{\Gamma}{\cU[\gamma] \equiv \cU }{j + 1} \quad
  \goodType{\Gamma}{\cB[\gamma] \equiv \cB}{0} \quad 
  \goodType{\Gamma}{\bot[\gamma] \equiv \bot}{0}
}
$$
\judgebox{ \goodTerm{\Gamma}{t}{T} }
$$
\Rule[]
{\goodType{\Gamma}{T}{j}}
{\goodTerm{\Gamma}{"c" \ T}{\cU}
}\quad
\Rule[]
{\goodTerm{\Gamma}{T}{\cU}}
{\goodType{\Gamma}{El \ T}{j}}
\quad
\Rule
{}
{\goodTerm{\Gamma}{"tt", "ff"}{\cB}}
\quad 
\Rule[name=Term Subst]
{\goodTerm{\Delta}{t}{T}
  \quad {\goodSub{\Gamma}{\gamma}{\Delta}}}
{\goodTerm{\Gamma}{t[\gamma]}{T[\gamma]}}
$$
$$
\Rule[]
{\goodTerm{\Gamma, A}{t}{B}}
{\goodTerm{\Gamma}{\lambda t}{\Pi A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{u}{A} 
\quad \goodTerm{\Gamma}{v}{B[(id, u)]}}
{\goodTerm{\Gamma}{(u,v)}{\Sigma A B}}
\quad 
\Rule[]
{\goodTerm{\Gamma}{t}{\Sigma A B}}
{\goodTerm{\Gamma}{"pjl" \ t}{A}
\quad \goodTerm{\Gamma}{"pjr" \  t}{B[(id, "pjl" \  t)]}
}
$$
$$
\Rule
{}
{\goodTerm{\Gamma}{(\lambda t)[\gamma] \equiv \lambda (t[\gamma^\uparrow])}{\Pi A B}
\quad \goodTerm{\Gamma}{(u,v)[\gamma] \equiv (u[\gamma],v[\gamma])}{\Sigma A B}
\quad \goodType{\Gamma}{El \ (T[\gamma]) \equiv (El \ T) [\gamma]}{}
}
$$

@@

We have shown some exemplar rules for all four standard types of
judgments. Now we focus on the newly introduced facility. We have two kinds of signatures, one for inductive types, the other for linkages.

@@

Recall in W-type~\cite{martin1982constructive}, a pair of type $x : A
\vdash B(x)$ decides a W-type. To simulate extensible inductive types, we
have to equip our inductive types with multiple constructors, and for
simplicity, each constructor has a fixed form \mintinline{agda}{cstrᵢ :
(x : Aᵢ) → (Bᵢ x → W) → W}.
Thus, we use a `list' of pairs ($\goodWSig{\Gamma}{w^+~A_i~B_i~\tau}{}$) as the signature of one
inductive type. Given an inductive type signature, we can use
\ruleref{WSig-Proj} to extract the corresponding $A_i, B_i$.

Of course we need to put an eye on the substitution law for each piece
of the syntax. Here $\bW \sigma$ is a large singleton
type~\cite{stone2000}, and it only has one ``element''---the inductive
type itself, inside. We formulate in this cumbersome way because we want
to have inductive type to be a field element and exposing the
constructors at the type level for later \textit{exhaustiveness checking} (Since otherwise, if an inductive type as field element has type
$\cU$, then no information about the inductive type is exposed).

@@

$\goodSig{\Gamma}{\sigma}{n}$ means $\sigma$ is a well-typed signature
for linkages in the context $\Gamma$ with $n$ fields. We omit the
naming/label information for fields in the meta\-theory. The signature (for the linkage) and the linkage rules are quite similar to
the signature for inductive types.  They are both constructed from the
empty stuff, and respect substitution lemmas. We have
\ruleref{Sig Proj}, \ruleref{Lnkg Proj} projection for each component.
We do have harmony~\cite{pfenning2009lecture} (i.e. local soundness
and local completeness, or $\beta,\eta$ rules). We omit them and the
substitution laws for either field addition and projection. For
$\mu^+,\nu^+$ rules, we will omit the sealing $\{f\}$ if possible.
% \footnote{This omission can be supported
%  by Agda using Implicit Arguments}

\textit{A side note}: before we dive in more details of our metatheory for linkage (especially the meaning of sealing and $\cC$), we have to mention that, the challenge of incorporating family polymorphism with inductive type mentioned in \ref{chg:extensible-inductive-type} mutates in metatheory because of de Bruijn indices. (Using named variables), consider the example of a linkage $\{a~:~\bW \sigma; b_1 : a \to T;b_2 ..\}$ with a bunch of $b_i$ \textit{using} $a$. The tension we need to solve here is between \textit{exhaustiveness checking} and \textit{inheritance (code reuse)}---on one hand, if we want to make sure $b_i$ is pattern matching correctly, then we have to construct $b_i$ term in the context of $a : \bW\sigma$; on the hand, if we want to let $b_i$ to be inherited into children family with extended inductive type $a~:~\bW \sigma'$, then the old inductive type $a~:~\bW \sigma$ in the context of $b_i$ is stopping that and thus we hope $b_i$ constructed in the context of $a : \cU$ (and thus later this $a$ can be compatible with inductive type of any signature). Our metatheory formulation of linkage is mainly trying to solve this tension.

We use $\cL$ to retrieve the corresponding types of linkage of a given signature, and $\cC$ to get the compiled linkage type (an existential type).
$\cL\sigma$ corresponds to the type of a family term, but $\cC\sigma$, according to \ruleref{Compile}, is a deeply nested dependent pair. $\cCt \sigma$ provide the concrete module of type $\cC\sigma$ given a linkage of type $\cL\sigma$.

The reason we need this additional type construct of $\cC\sigma$ here is
to support \textit{code reuse}, as the rule of linkages hinders its
ability to do abstraction. Concretely speaking, (using named variables)
we don't have \\ $\goodTerm{\Gamma, \cL(\{a~:~\bW \sigma; b~:~a \to
T\})}{h}{\cL(\{a~:~\cU; b~:~a \to T\})}$\footnote{If we had one,
consider the case $T$ is bottom type and $\sigma$ is empty record and
thus $a$ is initially bottom type. Now the direct projection using
\ruleref{Sig Proj} $\goodTerm{\Gamma, a : \cU}{h.b}{a \to \bot}$ will
lead to problem because every field in linkage has this implicit
universal quantification, and for this case we can substitute $a$ with
arbitrary type, say $\top$: $\goodTerm{\Gamma}{h.b[a \mapsto \top]}{\top
\to \bot}$ is a contradiction}.
However, this underivable term is important for us to abstract the
concrete inductive type so that the later field can depend on the
``abstract interface'' (i.e. something like $\cL(\{a:\cU~;~b:a \to T\})$) and achieve \textit{code reuse}. 
\YZ{It reads as if you want to have 'h : L\{a : U, b : a -> T\}'. But it is problematic because it leads to inconsistency, as you describe in a footnote.}\EDJreply{Good point. I rephrased it so the introduction of sigma type is smoother. Please check.}
\\ The good news, we don't have to stick with $\cL(\sigma)$ to express the abstract interface. We can also use sigma type, and actually 
we do have $\goodTerm{\Gamma, (\Sigma a : \bW\sigma, a \to T)
}{h}{(\Sigma a : \cU, a \to T)}$ because of the rules for sigma type.\footnote{The intro rule for sigma type decides that there is no such implicit universal quantification happening.}
\YZ{This sentence still speaks of overriding. Explain in terms of projection instead?}\EDJreply{Fixed. Please check.}
Thus to allow abstraction happens, we need $\cC\sigma$ to compile a
linkage to module as a part of the context and we also need
\ruleref{Lnkg Compile} to specify the compilation from linkage to
module.

This is also reflected in the implementation of our plugin : we know every field is defined in the context of the former fields, but our implementation actually make every field dependent on \textbf{the compiled module of the already-defined-part of the family}. 
%\YZ{Why is Csigma needed? How is it different from Lsigma?}\EDJreply{It corresponds to the compilation of the family ``type'', which is part of the implementation detail. Basically we know every field of a family is dependent on the former fields. The implementation detail here is that, every field of a family is dependent on the compiled module of the former family (and thus includes all the former fields)}\YZreply{I don't think you should explain the metatheory from the angle of how things are implemented in your plugin. Rather, when describing your Coq plugin, you can say it matches the metatheory.
%PS: you use two terminologies: 'compile' and 'seal', but I guess they mean very similar things. So avoid saying 'compile' here.
%PPS: Sealing means hiding, so what is being hided? Why do you need to hide it? And how does the 'Compile' and 'Seal' rules perform hiding?}\EDJreply{  Compile is very different from seal.  Compile is making a linkage into a module, depriving the ability to override---if we understand a linkage as a module with a lot of universal quantifier inside. Compile cannot hide things, but only translate a linkage into a module. Seal is really something that does hiding---because existential type can hide stuff. Seal is an term between existential types. \\
%The ultimate reason we have this two notion is that, the introduction rule for linkage fails to provide abstraction mechanism so we have to translate them into module. \\
%Generally speaking Compile : Linkage -> Module, Seal : Module -> Module, we need module here because linkage cannot do abstraction of inductive type. A concrete detail is at the above. \\ 
%I now describe the motivation of compile in the earlier paragraph, right before the "reflecting" section. Note that, this compile is inspired by plugin implementation. This is the reason I say the metatheory is really close and inspired by the implementation. Similarly as I say ``there is no extensible inductive type but only overriding of new inductive type'' because in metatheory it is using overriding to ``extend'' (same as plugin implementation), seal means the same thing, and ``compile'' in metatheory happens in plugin implementation
%}

@@

For notation clarity, we will sometimes use $"sf'"$ to emphasize
that we consider that seal $\goodSeal{\Gamma}{s}{\sigma}{A}$ as a term $\goodTerm{\Gamma, \cC \sigma}{sf'~s}{A[\pi_1]} $ and "sf" notation (used in \ruleref{Compile}) is defined as \\ $\goodSub{\Gamma, \cC \sigma}{(sf~s) := (\pi_1, sf'~s)}{\Gamma, A}$.\YZ{What is sf and sf'? sf first appears in Compile rules, so explain sf there.}\EDJreply{Fixed. They both are explained at here now}

Sealing judgement is still about the important term $h$ we mentioned
earlier. Thanks to $\cC$, a field $t_{new}$ that depends on $a$ is now
in the context containing $\cC\{a : \bW \sigma; ..\}$.
But if we want $t_{new}$ to be inherited(reused) when $\sigma$ is
extended, we would like $t_{new}$ to be in the context containing
$\cC\{a : \cU; ..\}$ instead. The sealing is responsible for this
because we have a sealing $\goodTerm{\Gamma, \cC\{a : \bW \sigma;
..\}}{s}{\cC\{a : \cU; ..\}}$ = $\goodTerm{\Gamma, (\Sigma a : \bW
\sigma,..)}{s}{(\Sigma a : \cU,..)}$. Then we can construct
$t_{new}$ in the context containing $(\Sigma a : \cU,..)$, ready for
inheritance. Thus sealing itself can be considered as proving 
abstractions on sigma type.

%\YZ{It occurs to me that you use the word "sealing" to mean
%different (though tangentially related) things. Previously you used it
%as in "sealed families".}\EDJreply{They are actually the same thing.
%This sealing in metatheory is much stronger than that of the implemented
%plugin. Downside of the strong sealing is that if the user really use
%this version of sealing, the user will have to annotate what the sealed
%family type is. Now in the plugin, the sealed family type is almost
%fixed (directly inferred using a fixed rules). So basically this
%complete version of sealing (in metatheory) can deduce that of seal (in
%plugin implementation)}

@@

Now, together with sealing, we can read \ruleref{Sig Add} and \ruleref{Lnkg Add} out clearly: given a constructed linkage $o : \cL\sigma$ of signature $\sigma$, the next added field $t : T$ will have to be in the context of the abtraction $A$ of the compilation $\cC\sigma$ (witnessed by the seal $f$). The ``concatenation'' of the $o$ and $t$ leads to a new linkage, of which the signature is the ``concatenation'' of the $\sigma$ and $T$ (we also record the sealing $f$ for simplicity). The key here is our signature, will \textit{record} this abstraction witness $f$, thus any projected member will still be in the abstracted context and thus possibly \textit{inheritable}.  The construction of linkage in \ruleref{Lnkg Add} requires $t$ to \textbf{only} refer to the (abstraction $A$ of) former part of the defining family $o : \cL\sigma$, this is to avoid inconsistency mentioned in \ref{chg:consistency} in metatheory.
